---
title: "README"
author: "Group 2"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Early-Stage Diabetes Risk Prediction Analysis

## Group 2

### Group Details

Group Leader: Yon Hui Yi 20022992

Group Members:\
1. Siti Sarah binti Mohd Azman 16100349\
2. Michelle Lee Yee Lin 22113922\
3. Adriel Soon Wei Chen 23065535\
4. Janice Beh Jing Ni 23064785

\

## README.md

The description of the dataset with the data structure and data
dictionary for the research project.

\

### Data Description

This dataset contains essential information on vital signs and symptoms
of individuals showing early signs of diabetes or those at risk of
developing diabetes. The variables in the dataset offer valuable
information about potential indicators of diabetes onset. It covers a
wide range of data, including demographic details and specific symptoms
associated with diabetes.

\

### Data Source

The research data was sourced from Kaggle, by the author
[ZJ](https://www.kaggle.com/datasets/tanshihjen/early-stage-diabetes-risk-prediction/data)
in November 2023

| **Observations** | **Variables** |
|:-----------------|:--------------|
| 520              | 17            |

\

## Data Dictionary

#### Research Data

All variables in the research dataset.

```{r}
library(kableExtra)

# Create the data dictionary
data_dict <- data.frame(
  Variables = c("Age", "Gender", "Polyuria", "Polydipsia", "Sudden Weight Loss", "Weakness", "Polyphagia", "Genital Thrush", "Visual Blurring", "Itching", "Irritability", "Delayed Healing", "Partial Paresis", "Muscle Stiffness", "Alopecia", "Obesity", "Class"),
 
   Description = c("Age range of the individuals.",
                  "Gender information.",
                  "Presence of excessive urination.",
                  "Excessive thirst.",
                  "Abrupt weight loss.",
                  "Generalized weakness.",
                  "Excessive hunger.",
                  "Presence of genital thrush.",
                  "Blurring of vision.",
                  "Presence of itching.",
                  "Display of irritability.",
                  "Delayed wound healing.",
                  "Partial loss of voluntary movement.",
                  "Presence of muscle stiffness.",
                  "Hair loss.",
                  "Presence of obesity.",
                  "Diabetes classification.")
)

# Print the data dictionary using kable
kable(data_dict, "html") %>%
  kable_styling()
```
